---
title: "Testing custom state object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{state}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(ppmcmc)
```

We can create a model state object in R that contains the state of each model
element (compartment) at a single time point.

```{r}
# Age groups
na <- 5
# Heterogeneity groups
nh <- 5
# New state
s <- state$new()
s$S <- matrix(0L, ncol = na, nrow = nh)
s$I <- matrix(0L, ncol = na, nrow = nh)
s$R <- matrix(0L, ncol = na, nrow = nh)
```

The state can have custom methods that under the hood call c++ functionality:

```{r}
s$I
s$update()
s$I
```

This maintains passing by reference, so is fast:
```{r}
s$I
system.time({
  for(i in 1:10000){
    s$update()
  }
})
s$I
```

For the pMCMC, we will need to be able to copy particles. To do this we need to 
force a copy of the state in memory, which can be specified with a custom deep clone method.

```{r}
# Without a deep clone, updating s, also changes s2
s2 <- s$clone()
s2$I
s$update()
s2$I

# With a deep clone it doesn't
s3 <- s$clone(deep = TRUE)
s3$I
s$update()
s3$I
```