---
title: "Testing speed of passing objects between R and c++"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing_speed}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(ppmcmc)
library(microbenchmark)
library(ggplot2)
```

Here we explore passing objects efficiently between R and c++ (using Rcpp). This will
likely be obvious to a computer scientist but...let's learn by doing.

Firstly we have written two function in c++ that take a matrix and modify it by 
adding a constant. This is either done:
1. Across a row, using `add_to_row()` or
2. Down a column using `add_to_col()`.
Rcpp passes by reference by default and therefore these functions modify in place 
(unless a copy of the matrix passed to them is forced to be made).

To show what these functions do:
```{r}
x <- matrix(0L, nrow = 5, ncol = 5)
x

add_to_row(x, 0, 1)
x

add_to_col(x, 0, 1)
x
```

We can test the efficiency impact of modifying row-wise or column-wise:

```{r, message=FALSE, fig.width=10}
test_row <- function(n){
  x <- matrix(0L, nrow = n, ncol = n)
  add_to_row(x, 0, 1)
}
test_col <- function(n){
  x <- matrix(0L, nrow = n, ncol = n)
  add_to_col(x, 0, 1)
}
row_v_col <- microbenchmark(test_row(1000), test_col(1000), times = 100)
ggplot2::autoplot(row_v_col)
```
We can see that, as the  elements of the matrix are stored in the vector in memory
in column-major order, operating column-wise is quicker. However the differences
are quite small (~ 100 microseconds, or 1 ten thousandth of a second).


We need to be much more careful about making copies. We can show that is more 
inefficient if a copy is forced to be made if we pass a numeric matrix (not integer):

```{r, message=FALSE, fig.width=10}
test_col_copy <- function(n){
  x <- matrix(0, nrow = n, ncol = n)
  add_to_col(x, 0, 1)
}
col_v_copy <- microbenchmark(test_col(1000), test_col_copy(1000), times = 100)
ggplot2::autoplot(col_v_copy)
```

Lets show how quickly we can do many steps between R and c++, modifying a large
matrix at each step.
```{r, message=FALSE, fig.width=10}
pass_back_and_forth <- function(x){
  out <- rep(0, 1000)
  for(i in 1:1000){
    add_to_col(x, 0, i)
    out[i] <- x[1,1]
  }
  return(out)
}
x <- matrix(0L, nrow = 1000, ncol = 1000)
out <- pass_back_and_forth(x)
head(out)
ggplot2::autoplot(microbenchmark(pass_back_and_forth(x), times = 10))
```


